Real-space computation of $E$/$B$-mode maps I: Formalism and Compact Kernels

### 2018-06-11 articles/formalism/sections/discussion.tex
added explanation for the alpha and gamma as being the radiating and the convolution kernels, still sharpening the explanation
Editing pass and notation cleanup.
Added some reference and some detail to kernel paintbrush section.

In this article we have cast the standard CMB polarization analysis operations in a vector matrix notation. Using this concise notation we derive the real space operators that translate the Stokes vector \vp{} to  the vector of scalars \vs and vice versa. We explicitly demonstrated that this real space operation can be simply interpreted as a convolution over the complex field $[Q - i U]$ (or $[E + iB]$) with an effective complex beam which is fully expressed in terms of the $Y_{\ell 2}$ spherical harmonic functions. We also use this vector matrix notation to derive real space operators which allow the direct decomposition of the full Stokes vector \vp{} into the vector \vp{E} and \vp{B} that correspond to the respective scalar modes.

Given the effective beam interpretation of these real space operators we derive the harmonic coefficients of these effective beams at the north galactic pole. Using these harmonic coefficients we provide a prescription for computing the convolution kernels at any position on the sphere using the standard Healpix built in functions. The procedure is equivalent to parallel transporting the beam at the north pole to any desired location on the sphere. We implement the prescription to compute the  kernel at different location on the sphere and provide simple explanations in terms of Euler angles for the observed variations.

These real space convolution kernels provide a spatially intuitive way of understanding the construction of the scalar modes. We explicitly show that the kernels separates into an band limit independent azimuthal operation around any given direction which is primarily responsible for requisite decomposition, while the band limit dependent radial weights can be interpreted as some isotropic smoothing operation. These radial weights primarily determine the non-local dependence of the construction of the respective fields at any location on the Stoke field. We define the parameter $\beta_0$ as a means to characterize the non-locality and show that $\beta_0$ scales $\propto \ell_{\rm max}^{-1}$. We show that this non-locality parameter also characterized the non-locality of the $\bar{O}_{E/B}$ operators.

@@

>>>>

\begin{itemize}
\item Summarize what has been achieved in this peice of work.
\item Whats new as compared to work done by Zaldariaga?
\item Not having to compute spin spherical harmonics.
\item Potentially allows for TOD to E/B map generation.
\item Compact kernels and understanding of why mask apodization works.
\item Can you deconvolve anisotropic compact kernels? Meaning different radial cut-off in different portions of the sky. This will be useful to dealing with foregrounds. Its not obvious if this can be achieved in harmonic space.  
\item The relation to galaxy shear E/B estimates.
\item Understanding E/B signatures of magentized filaments.
\end{itemize}

\revisit{A similar equation for real space $E$ \& $B$ operators was derived in \cite{Zaldarriaga2001a}, however those results were derived for the flat sky case and did not explicitly derive the radial kernel.} \comment{A discussion on this should be in the conclusions.}

In this work we present real space operators on the sphere that transform maps of Stoke Q/U parameters to scalars E/B. We  also present real space operators that decompose maps of Stokes Q/U parameters into the parts that only contribute to E/B modes. We introduced a vector-matrix notation which allows for concise book keeping of all the operations involved and simplifies the derivation of the real space operators.

These real space operators provide a spatially intuitive way of understanding construction of the scalar modes. We explicitly  demonstrated that all the operators can be separated into a band limit independent azimuthal operation and band limit dependent radial weights. The azimuthal part of the operator is primarily responsible for the requisite spinorial decomposition, while the radial weights determine the non-local dependence of the construction of the resultant fields. We define the $\beta_0$ parameter using an empirical relation to characterize the non-locality and demonstrate that it scales $\propto \ell_{\rm max}^{-1}$.

Our careful study of the real space operators reveals the dual interpretation of the operators as either the Greens function or a convolving beam depending on whether the kernels are expressed in terms of the forward rotation Euler angles or the inverse rotation Euler angles. While the convolution interpretation is familiar to all, the Green's function interpretation is a new one arising from this work. In particular the Green's function interpretation allows us to think of ${}_{+2}X$ as some spin-2 charge which radiates out a complex spin-0 scalar field $E+iB$. The resultant E/B maps can be then understood as arising from superposition of the radiating field from all the spin charges on the sphere.

 We explicitly demonstrated that this real space operation can be simply interpreted as a convolution over the complex field $[Q - i U]$ (or $[E + iB]$) with an effective complex beam which is fully expressed in terms of the $Y_{\ell 2}$ spherical harmonic functions. We also use this vector matrix notation to derive real space operators which allow the direct decomposition of the full Stokes vector \vp{} into the vector \vp{E} and \vp{B} that correspond to the respective scalar modes.

@@

\input{sections/kernel_paintbrush.tex}

### 2018-06-11 articles/formalism/sections/generalizedeb.tex

%--------------------------------------------------------
%--------------------------------------------------------
\subsection{Generalized polarization operators}
With our better understanding of the radial part of the kernel for CMB polarization, we can write down generalized $E/B$-like fields that depend on a different radial function, even one that we specify to have compact support.
The spin symmetry constrains the the azimuthal part of the real space kernels to be of the form $\sim e^{\pm i2 \alpha}$.  The shape of the radial part of the operator is determined by the basis functions and varies as a function of the band limit. It is here that we may potentially choose alternate forms for the radial functions to suit certain kind of analysis.

>>>>

%--------------------------------------------------------
%--------------------------------------------------------
\subsection{Generalized polarization operators}\label{sec:generalized_operators}
With our better understanding of the radial part of the kernel for CMB polarization, we can write down generalized $E/B$-like fields that depend on a different radial function, even one that we specify to have compact support.
The spin symmetry constrains the the azimuthal part of the real space kernels to be of the form $\sim e^{\pm i2 \alpha}$.  The shape of the radial part of the operator is determined by the basis functions and varies as a function of the band limit. It is here that we may potentially choose alternate forms for the radial functions to suit certain kind of analysis.

### 2018-06-11 articles/formalism/sections/introduction.tex

This paper is organized in the following manner: In \sec{sec:pol-primer} we present a primer on the description of CMB polarization on the sphere. Here we begin with a heuristic argument that makes transparent the real space construction of E/B modes on the sphere. We then discuss the standard harmonic space procedures for this operation. Finally we introduce the matrix-vector notation which yields a more concise description of the harmonic space procedures. In \sec{sec:real_space_operators} we derive and discuss the real space operators that transform Q/U to E/B and vice versa. In \sec{sec:visualize_operator} we evaluate these real space operators and present visualizations of these functions. We also derive a real space operator that decomposes the Stokes Q/U  parameters into components that correspond to E and B modes respectively.  Here we also discuss the locality of the real space E \& B operators. In \sec{sec:numerical_implementation} we implement these operators to evaluate E \& B  maps from the Stokes parameters Q \& U and compare these maps and their spectra from those derived using Healpix. We conclude with a discussion and the scope of this new method of analyzing CMB polarization in \sec{sec:discussion}.

>>>>

This paper is organized in the following manner: In \sec{sec:pol-primer} we present a primer on the description of CMB polarization on the sphere. Here we begin with a heuristic argument that makes transparent the real space construction of E/B modes on the sphere. We discuss the standard harmonic space procedures for this operation. Finally we introduce the matrix-vector notation which yields a more concise description of the harmonic space procedures. In \sec{sec:qu2eb} and \sec{sec:eb2qu} we derive and discuss the real space operators that transform Q/U to E/B and vice versa. In \sec{sec:visualize_operator} we evaluate these real space operators and present visualizations of these functions. In \sec{sec:purify_stokes_qu} we derive a real space operator that decomposes the Stokes Q/U  parameters into components that correspond to E and B modes respectively and present its visualizations.  In \sec{sec:radial_locality} we study the locality of the real space operators and explore its band limit dependence. In \sec{sec:generalized_operators} we present a systematic way to generalize the real space operators by varying the non-locality parameter and discuss the connection to the spin raising and lower operators. We conclude with a summary and discuss the scope of this new method of analyzing CMB polarization in \sec{sec:discussion}.

### 2018-06-11 articles/formalism/sections/kernel_discussion.tex

To quantify the non-locality of the scalar modes $E/B$, we can define a characteristic angular radius of the region from which the kernels get most of their contribution.   We define a non-locality parameter $\beta_{0}$ as the angle beyond which the function ${_{\mm}f}(\beta,\ell_{\rm min}=2,\ell_{\rm max})$ consistently falls below 1\% of its maximum.
%For $\ell_{\rm max}=24$, the maximum multipole accessible on a Nside=8 Healpix map, the non-locality parameter $\beta_0=180^{\circ}$ as the radial function never falls monotonously below 1\% of its global maxima as seen in \fig{}. Using this fact and the self similar property of the radial functions, we define the following empirical relation: $\beta_o= {\rm min}(180,180 \frac{24}{\ell_{\rm max}})$, as a means of estimating the non-locality parameter given the maximum multipole $\ell_{\rm max}$ accessible for analysis.
The empirical relation
\beq
\beta_0= {\rm min}\left(180,180 \frac{22}{\ell_{\rm max}} \right)
\eeq
provides a reasonable estimate of this transition point for  ${}_{\mm}f$. \revisit{This empirical function also provides a reasonable estimate for the transition point for the function ${}_{\md}f$ particularly for high $\ell_{\rm max}$ and it systematically over-predicts the transition point for the function ${}_{\mi}f$ as seen in \fig{fig:rad_ker_decay}.}

>>>>

To quantify the non-locality of the scalar modes $E/B$, we can define a characteristic angular radius of the region from which the kernels get most of their contribution.   We define a non-locality parameter $\beta_{0}$ as the angle beyond which the function ${_{\mm}f}(\beta,\ell_{\rm min}=2,\ell_{\rm max})$ consistently falls below 1\% of its maximum.
%For $\ell_{\rm max}=24$, the maximum multipole accessible on a Nside=8 Healpix map, the non-locality parameter $\beta_0=180^{\circ}$ as the radial function never falls monotonously below 1\% of its global maxima as seen in \fig{}. Using this fact and the self similar property of the radial functions, we define the following empirical relation: $\beta_o= {\rm min}(180,180 \frac{24}{\ell_{\rm max}})$, as a means of estimating the non-locality parameter given the maximum multipole $\ell_{\rm max}$ accessible for analysis.
\revisit{The empirical relation:
\beq
\beta_0= {\rm min}\left(180,180 \frac{\ell_{0}}{\ell_{\rm max}} \right) \,,
\eeq
with $\ell_{0}=22$ provides a reasonable estimate of this transition point for ${}_{\mm}f$ as seen in \fig{fig:rad_ker_decay}. Setting $\ell_{0}=10$ and $\ell_{0}=32$ provides the transition points for the functions ${}_{\mi}f$ and  ${}_{\md}f$ respectively.}
